---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Some Information
The variables included in the Activity.csv dataset are:

- steps: Number of steps taking in a 5-minute interval (missing values are coded as NA) 

- date: The date on which the measurement was taken in YYYY-MM-DD format

- interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

## Loading and preprocessing the data
```{r}
## Load the data (i.e. read.csv())
data <- read.csv('activity.csv')
cat("Data Dimensions: ", dim(data), "\n\n")
cat("Data Summary: \n", summary(data), "\n\n")
head(data)

## Process/transform the data (if necessary) into a format suitable for your analysis
data_no_na <- na.omit(data)
```


## What is mean total number of steps taken per day?
```{r}
## For this part of the assignment, you can ignore the missing values in the dataset.

## 1. Make a histogram of the total number of steps taken each day
### Create table calculating total steps
total_steps <- aggregate(data_no_na$steps, by=list(data_no_na$date), sum)
### Rename columns (makes it easier to make histogram)
colnames(total_steps)[1:2] <- c("date", "total_steps")
### Create histogram
hist(total_steps$total_steps, main ="Histogram of the Total Number of Steps Taken Each Day", xlab = "Total Number of Steps")

## 2. Calculate and report the mean and median total number of steps taken per day
### Caluclate mean
mean <- mean(total_steps$total_steps)
### Calculate median
median <- median(total_steps$total_steps)
### Print results
cat("Mean: ", mean, "\n\n")
cat("Median:", median)
```

## What is the average daily activity pattern?
```{r}
## 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number 
##    of steps taken, averaged across all days (y-axis)

## 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```

## Imputing missing values
```{r}
## Note that there are a number of days/intervals where there are missing values (coded as NA). 
## The presence of missing days may introduce bias into some calculations or summaries of the data.

## 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

## 2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. 
##    For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

## 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

## 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps 
##    taken per day. Do these values differ from the estimates from the first part of the assignment? 
##    What is the impact of imputing missing data on the estimates of the total daily number of steps?
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
## 1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date 
##    is a weekday or weekend day.

## 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
##    and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
##    The plot should look something like the following, which was created using simulated data:
```